\iflater
\section*{Budget Justification}

\todo{GPG: Budgets for all projects must include funding for one or
  more designated project representatives (PI/co-PI/senior researcher
  or NSF-approved replacement) to attend each annual PI meeting during
  the proposed lifetime of the award.}

\todo{GPG: a portion of the budget for each Core Programs, Large
  Projects proposal may be used to engage relevant Broadening
  Participation in Computing (BPC) expertise to help plan, organize,
  coordinate and execute BPC activities.}

\todolater{Make sure to account for PhD months accurately---note when
  people start.}

\hg{Any chance we want to budget for a new machine for someone (e.g., one or
more of the potential other PhD students)? Or is that priced in?}
\bcp{Good catch -- yes, we want to budget for new machines for all.
  Possibly also a server, for running lots of tests faster...}

\todo{REPL students (see comments here)}
% And attached is the budget sheet from PEFS. The TLDR yearly costs are….

% Direct Student Costs
% ================
% Travel - $400 per student per year
% Housing - $2500 per student per year
% Stipend - $6000 per student per year

% Indirect Student Costs
% =================
% Admin - $32,500 per year (10k for lead admin, 10k distributed across PLClub+ leaders, with Penn’s 62.5% overhead stacked on top)

% Let me know if you have any questions. If you want to contribute to indirect student costs, I’m not totally sure what the right amount is — but it seems we’re currently charging about $4,000 per student per year in admin costs.

\discuss{Should we be asking for budget for a compute server or AWS or
  any other Cloud Computing service (e.g., NSF itself has one)), for
  running heavy testing workloads??  If not, we should say so.  The
  benchmarking project, for example, could use this!}

% No more than 3 pages!!!
\subsection*{A. Senior Personnel}
\noindent{\bf A1.} Includes two PIs at 18\% CY.
\subsection*{B. Other Personnel}
\noindent{\bf B3.} Includes stipend for one graduate student for each calendar year of the project.

\todo{Webinar: Staff support is normal.  Needs to justified not only
  in the body of the proposal but also carefully called out in the
  budget justification.}

\subsection*{C. Fringe Benefits}
Fringe benefits are calculated at a rate of X\% for faculty, Y\% for graduate students.
\subsection*{E. Travel}
1) all travel (both domestic and foreign) must now be justified.
2) temporary dependent care costs above and beyond regular dependent care that directly result from travel to conferences are allowable costs provided that the conditions established in 2 CFR § 200.474 are met.
\subsection*{G. Other Direct Costs}
1) Includes coverage on costs of computing devices
2) The charging of computing devices as a direct cost is allowable for devices that are essential and allocable, but not solely dedicated, to the performance of the NSF award
\noindent{\bf G5.} Includes tuition for graduate students participating in the program.
\subsection*{H. Indirect Costs}
Overhead at a rate of X\% is charged on all direct salaries and wages, applicable fringe benefits, materials and supplies, services, travel and subawards up to the first \$X of each subaward. Excluded are equipment and the portion of each subaward in excess of \$X.
\fi
