\section*{Project Summary}

\paragraph*{Overview}
Property-based testing (PBT) is an advanced software testing
methodology where users write executable specifications of their code
and an automated test harness checks that they hold for many randomly
generated test inputs.  Its bug-finding power arises from its ability
to exercise a wide range of system behaviors---both expected and
unexpected---with minimal user guidance.
%
From its roots in the QuickCheck library in Haskell, PBT has grown
into a standard tool across the functional programming community, and it
has begun to make inroads into industrial practice
% and it is more recently being incorporated into testing processes
in companies such as Quviq, Galois, and Amazon.\iflater\todo{Who else??}\fi{}
% ... where it finds lots of bugs!
%
The goal of this project is to accelerate this transition
by identifying and addressing key challenges in
present-day PBT tools.

What are these challenges?
A need-finding study of PBT at Jane
Street, a Wall Street firm with a focus on software
technology as a competitive advantage, found developers enthusiastic
about its {\em usefulness} but
sometimes frustrated by issues with {\em usability}.
%
The study identified usability issues in both of the major components
of PBT tooling---{\em specification} of properties and {\em
  generation} of random inputs. Furthermore, it identified challenges
around another aspect of PBT that has received less attention in the
research literature: how programmers can {\em validate} the
effectiveness of testing.

% At the level of {\em specification}, developers wished for better
% support for expressing common requirements such as ``this
% high-performance implementation of a service must behave exactly the
% same as that reference model'' and for articulating properties of
% poorly modularized code.
% %
% Concerning the other main component of PBT, the {\em generation} of
% random test cases, \todo{...}
% %
% Developers also need better {\em validation} tools for checking
% that their testing is actually effective.\todo{blah}
% %
% Finally, better {\em education} of potential PBT users is critically
% required.

% \amh{How about the following framing? In this project, we define a practice of
% usable property-based testing, and conduct foundational research in PL and HCI
% to bring about this practice. We define usable property based-testing as having
% two main components. First, providing visibility into what can be extremely
% complex test outcomes. Second, supporting the efficient expression of
% tests.\ldots{} And then we can jump into talking about the challenges that we
% plan to solve in specification, generation, validation, and education.}.
% %
% At the level of {\em specification}, developers wished for better
% support for expressing common requirements such as ``this
% high-performance implementation of a service must behave exactly the
% same as that reference model'' and for articulating properties of
% poorly modularized code.
% %
% Concerning the other main component of PBT, the {\em generation} of
% random test cases, \todo{...}
% %
% Developers also need better {\em validation} tools for checking
% that their testing is actually effective.\todo{blah}
% %
% Finally, better {\em education} of potential PBT users is critically
% required.

%  and ``shrinking'' of failing
% tests to minimal counterexamples,

To make progress, two distinct---and largely disconnected---research areas
must be brought to bear.  PBT itself is grounded in domain-specific
languages and formal methods: traditional PL topics.  Usability, on
the other hand, is the domain of human-computer interaction.
%
This four-year project will combine insights, methods, and tools from
PL and HCI to advance the state of the art in all three components of
PBT---specification, generation, and validation---to make PBT
significantly more usable across a broad community of software developers.

\smallskip

\noindent{\bf Keywords:} Property-based testing, human-computer
interaction, programming languages, usability.

\paragraph*{Intellectual Merit}
The project will advance knowledge along five interconnected axes.
%
First, it will establish a firm {\bf\em foundation} for HCI-informed
research on PBT, supplementing past and ongoing user studies with
broader surveys of PBT across the software industry and real-time
observations of developers using PBT tools.
%
Second, it will offer developers more usable {\bf\em specification} tools,
including a language for stating temporal properties over internal
program states and a tool for automating model-based testing of
modular abstractions.
%
Third, it will explore a novel abstraction for random input {\bf\em
  generation} that enables a range of use cases---generating inputs
satisfying validity conditions, mutating inputs to explore the space
of similar inputs, and manually or automatically tuning a random
generator's distribution based on examples or on code coverage
measurements.
%
And fourth, it will build new interactive programming tools for rapid
{\bf\em validation} of the effectiveness of testing, helping programmers
pin down the causes of failures, and visualize generated data
distributions to support comprehension and tuning.

%  (1) deploy tools from HCI to better
% understand the challenges to wider adoption of PBT, (2) combine PL and
% HCI methods to build solutions, and (3) develop educational materials
% to make PBT concepts and tools accessible to a broad community of
% university students and industrial developers.

\paragraph*{Broader Impacts}
A fifth thread, coordinated with the rest and integral to the
project's aims, will be to develop materials for PBT {\bf\em education}
in both universities and industry, including an introduction to PBT for
an undergraduate data structures course, a tool to help students
author properties, and advice on how programmers can identify
high-leverage situations where properties are easy to find.

The ultimate goal, through both education and intuitive, freely
distributed tools, is to make property-based testing a standard option
on every software developer's testing toolbelt.  Better testing, in
turn, will lead to software systems of every sort that are more
robust, more reliable, and less expensive to build.

The work we do will include and elevate undergraduate researchers, including
many from diverse backgrounds that will be funded through a recently funded REU
program. These undergraduate students will work closely with PIs and PhD
students in an interconnected and supportive research environment.
\bcp{What else?  Working with undergrads, etc., etc.}
\hg{How's that? More?}

% The statement on broader impacts should describe the potential of the proposed activity to benefit society and contribute to the achievement of specific, desired societal outcomes.
